\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2021-09-07 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
fixed indexing issues with equations (7) and (8)

\subsubsection{Notation and organization}
%We use boldfaced small letters for vectors, and boldfaced capital letters for matrices. 
The set of natural numbers, real numbers and complex numbers are denoted as $\mathbb{N}$, $\mathbb{R}$, and $\mathbb{C}$, respectively. The symbol $\nabla_{\bm{x}}$ denotes the Euclidean gradient operator with respect to (w.r.t.) the vector $\bm{x}$. Thus, $\nabla_{\bm{x}}\cdot$ stands for the divergence, and $\Delta_{\bm{x}}$ stands for the Laplacian w.r.t. vector $\bm{x}$. We use $\langle\cdot,\cdot\rangle$ to denote the standard Euclidean inner product. The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively. We use the superscript $*$ to denote the complex conjugate. The uniform probability distribution over an interval $[a,b]$ is denoted as ${\rm{Unif}}\left([a,b]\right)$. Likewise, the $n$ dimensional uniform probability distribution over $[a,b]^{n}$ is denoted as ${\rm{Unif}}\left([a,b]^{n}\right)$. The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute magnitude, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\odot$ and $\oslash$ respectively denote the elementwise (Hadamard) product and division, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map. The $n\times n$ identity and zero matrices are denoted as $\bm{I}_{n}$ and $\bm{0}_{n\times n}$, respectively.

@@

With the $i$th generator, we associate its inertia $m_{i}>0$ and damping coefficient $\gamma_{i}>0$. The other parameters: the \emph{effective power input} $P_{i}$, the \emph{phase shift} $\varphi_{ij}\in[0,\frac{\pi}{2})$, and the \emph{coupling coefficients} $k_{ij}\geq 0$, depend on the network reduced admittance matrix $\bm{Y}\equiv[Y_{ij}]_{i,j=1}^{n}\in\mathbb{C}^{n\times n}$. Specifically,
\begin{subequations}
\begin{align}
 P_{i} &= P_{i}^{\text{mech}} - P_{i}^{\text{load}} - |E_{i}|^{2} \Re\left(Y_{ii}\right) + \Re\left(E_i \cdot I_{i}^{*}\right),\label{defPi}\\
 \varphi_{ij} &= \begin{cases} -\arctan\left(\dfrac{\Re\left(Y_{ij}\right)}{\Im\left(Y_{ij}\right)}\right), & \text{if}\;i\neq j,\\
 0, & \text{otherwise},
 \end{cases}\\
 k_{ij} &= \begin{cases} |E_{i}| |E_{j}| \vert Y_{ij}\vert, & \text{if}\;i\neq j,\\
 0, & \text{otherwise},
 \end{cases}
\end{align}
\label{DefParam}
\end{subequations}
where $P_{i}^{\text{mech}}$ is the mechanical power input, $P_{i}^{\text{load}}$ is the real load, $E_{i}$ is the internal voltage, and $I_{i}\in\mathbb{C}$ is the current for generator $i$.

@@

We define the \emph{positive diagonal} matrices
\begin{align*}
\bm{M} &:= {\rm{diag}}\left(m_{1},\hdots,m_{n}\right), \nonumber\\
\bm{\Gamma} &:= {\rm{diag}}\left(\gamma_{1},\hdots,\gamma_{n}\right),\nonumber\\
\bm{\Sigma} &:= {\rm{diag}}\left(\sigma_{1},\hdots,\sigma_{n}\right),
%\label{PosDiagMatrices}	
\end{align*}
and rewrite (\ref{ItoSDEcomponentlevel}) as a mixed conservative-dissipative SDE in state vector $\bm{x} := (\bm{\theta},\bm{\omega})^{\top} \in \mathbb{T}^{n} \times \mathbb{R}^{n}$ as
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\theta}\\
{\mathrm{d}}\bm{\omega}	
\end{pmatrix}
\! = \!\begin{pmatrix}
\bm{\omega}\\
-\bm{M}^{-1}\nabla_{\bm{\theta}}V(\bm{\theta}) -\bm{M}^{-1}\bm{\Gamma}\bm{\omega}  	
\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
 \bm{0}_{n\times n}\\
 \bm{M}^{-1}\bm{\Sigma}	
 \end{pmatrix}{\mathrm{d}}\bm{w},
\label{ItoSDEvectorlevel}
\end{align}
where 
%$\odot$ denotes element-wise product, $\oslash$ denotes element-wise division, the , the rotor inertia vector $\bm{m} = (m_{1},\hdots,m_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the rotor damping coefficient vector $\bm{\gamma} = (\gamma_{1}, \hdots, \gamma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the diffusion coefficient vector $\bm{\sigma} = (\sigma_{1}, \hdots, \sigma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, and 
$\bm{w}\in\mathbb{R}^{n}$ is the standard vector Wiener process, $\mathbb{T}^{n}$ denotes the $n$-torus $[0,2\pi)^{n}$, and the potential function $V : \mathbb{T}^{n} \mapsto \mathbb{R}$ is given by
\begin{align}
V(\bm{\theta}) := -\displaystyle\sum_{i=1}^{n} P_{i}\theta_{i} + \!\displaystyle\sum_{i,j=1}^{n}\!k_{ij}\left(1 - \cos(\theta_{i}-\theta_{j}-\varphi_{ij})\right).
\label{potential}	
\end{align}
%wherein $\mathcal{E}$ is the set of edges for the underlying graph modeling the power system network topology. 
The potential (\ref{potential}) has a natural energy function interpretation and can also be motivated by a mechanical mass-spring-damper analogy \cite{dorfler2013synchronization,ishizaki2018}. %The function $V$ is continuously differentiable, bounded below, and goes to $+\infty$ as $\|\bm{\theta}\|_{2}\rightarrow +\infty$.
%In the development below, an important role will be played by the ``Hamiltonian-like" function
%\begin{align}
%H(\bm{x})\equiv H(\bm{\theta},\bm{\omega}) := \frac{1}{2}\bm{\omega}^{\top}{\rm{diag}}\left(\bm{\gamma}\oslash\bm{m}\right)\bm{\omega} \: + \: V(\bm{\theta}).
%\label{hamiltonian}	
%\end{align}

>>>>

\subsubsection{Notation and organization}
%We use boldfaced small letters for vectors, and boldfaced capital letters for matrices. 
The set of natural numbers, real numbers and complex numbers are denoted as $\mathbb{N}$, $\mathbb{R}$, and $\mathbb{C}$, respectively. The symbol $\nabla_{\bm{x}}$ denotes the Euclidean gradient operator with respect to (w.r.t.) the vector $\bm{x}$. Thus, $\nabla_{\bm{x}}\cdot$ stands for the divergence, and $\Delta_{\bm{x}}$ stands for the Laplacian w.r.t. vector $\bm{x}$. We use $\langle\cdot,\cdot\rangle$ to denote the standard Euclidean inner product. The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively. We use the superscript $*$ to denote the complex conjugate, and the superscript $^{\top}$ to denote the matrix transpose. The uniform probability distribution over an interval $[a,b]$ is denoted as ${\rm{Unif}}\left([a,b]\right)$. Likewise, the $n$ dimensional uniform probability distribution over $[a,b]^{n}$ is denoted as ${\rm{Unif}}\left([a,b]^{n}\right)$. The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute magnitude, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\odot$ and $\oslash$ respectively denote the elementwise (Hadamard) product and division, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map. The $n\times n$ identity and zero matrices are denoted as $\bm{I}_{n}$ and $\bm{0}_{n\times n}$, respectively.

@@

With the $i$th generator, we associate its inertia $m_{i}>0$ and damping coefficient $\gamma_{i}>0$. The other parameters: the \emph{effective power input} $P_{i}$, the \emph{phase shift} $\varphi_{ij}\in[0,\frac{\pi}{2})$, and the \emph{coupling coefficients} $k_{ij}\geq 0$, depend on the network reduced admittance matrix $\bm{Y}\equiv[Y_{ij}]_{i,j=1}^{n}\in\mathbb{C}^{n\times n}$, $\bm{Y}=\bm{Y}^{\top}$. Specifically,
\begin{subequations}
\begin{align}
 P_{i} &= P_{i}^{\text{mech}} - P_{i}^{\text{load}} - |E_{i}|^{2} \Re\left(Y_{ii}\right) + \Re\left(E_i \cdot I_{i}^{*}\right),\label{defPi}\\
 \varphi_{ij} &= \begin{cases} -\arctan\left(\dfrac{\Re\left(Y_{ij}\right)}{\Im\left(Y_{ij}\right)}\right), & \text{if}\;i\neq j,\\
 0, & \text{otherwise},
 \end{cases}\\
 k_{ij} &= \begin{cases} |E_{i}| |E_{j}| \vert Y_{ij}\vert, & \text{if}\;i\neq j,\\
 0, & \text{otherwise},
 \end{cases}
\end{align}
\label{DefParam}
\end{subequations}
where $P_{i}^{\text{mech}}$ is the mechanical power input, $P_{i}^{\text{load}}$ is the real load, $E_{i}$ is the internal voltage, and $I_{i}\in\mathbb{C}$ is the current for generator $i$.

@@

We define the \emph{positive diagonal} matrices
\begin{align*}
\bm{M} &:= {\rm{diag}}\left(m_{1},\hdots,m_{n}\right), \nonumber\\
\bm{\Gamma} &:= {\rm{diag}}\left(\gamma_{1},\hdots,\gamma_{n}\right),\nonumber\\
\bm{\Sigma} &:= {\rm{diag}}\left(\sigma_{1},\hdots,\sigma_{n}\right),
%\label{PosDiagMatrices}	
\end{align*}
and rewrite (\ref{ItoSDEcomponentlevel}) as a mixed conservative-dissipative SDE in state vector $\bm{x} := (\bm{\theta},\bm{\omega})^{\top} \in \mathbb{T}^{n} \times \mathbb{R}^{n}$ as
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\theta}\\
{\mathrm{d}}\bm{\omega}	
\end{pmatrix}
\! = \!\begin{pmatrix}
\bm{\omega}\\
-\bm{M}^{-1}\nabla_{\bm{\theta}}V(\bm{\theta}) -\bm{M}^{-1}\bm{\Gamma}\bm{\omega}  	
\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
 \bm{0}_{n\times n}\\
 \bm{M}^{-1}\bm{\Sigma}	
 \end{pmatrix}{\mathrm{d}}\bm{w},
\label{ItoSDEvectorlevel}
\end{align}
where 
%$\odot$ denotes element-wise product, $\oslash$ denotes element-wise division, the , the rotor inertia vector $\bm{m} = (m_{1},\hdots,m_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the rotor damping coefficient vector $\bm{\gamma} = (\gamma_{1}, \hdots, \gamma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the diffusion coefficient vector $\bm{\sigma} = (\sigma_{1}, \hdots, \sigma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, and 
$\bm{w}\in\mathbb{R}^{n}$ is the standard vector Wiener process, $\mathbb{T}^{n}$ denotes the $n$-torus $[0,2\pi)^{n}$, and the potential function $V : \mathbb{T}^{n} \mapsto \mathbb{R}$ is given by
\begin{align}
V(\bm{\theta}) := -\displaystyle\sum_{i=1}^{n} P_{i}\theta_{i} + \!\displaystyle\sum_{\substack{i,j=1\\i<j}}^{n}\!k_{ij}\left(1 - \cos(\theta_{i}-\theta_{j}-\varphi_{ij})\right).
\label{potential}	
\end{align}
%wherein $\mathcal{E}$ is the set of edges for the underlying graph modeling the power system network topology. 
The potential (\ref{potential}) has a natural energy function interpretation and can also be motivated by a mechanical mass-spring-damper analogy \cite{dorfler2013synchronization,ishizaki2018}. %The function $V$ is continuously differentiable, bounded below, and goes to $+\infty$ as $\|\bm{\theta}\|_{2}\rightarrow +\infty$.
%In the development below, an important role will be played by the ``Hamiltonian-like" function
%\begin{align}
%H(\bm{x})\equiv H(\bm{\theta},\bm{\omega}) := \frac{1}{2}\bm{\omega}^{\top}{\rm{diag}}\left(\bm{\gamma}\oslash\bm{m}\right)\bm{\omega} \: + \: V(\bm{\theta}).
%\label{hamiltonian}	
%\end{align}
